---
title: "Microbiome Sequence Analysis"
subtitle: "Importing 16S Sequence Data"
author: "Sarah Tanja"
format: html
editor: visual
toc: true
toc-title: Contents <i class="bi bi-bookmark-heart"></i>
toc-depth: 5
toc-location: left
reference-location: margin
citation-location: margin
bibliography: ./references/qiime2.bib
---

## Overview

The 16S sequences were provided to me from Mr. DNA via a DropBox download link. They are **Demultiplexed** (aka **Demuxed**) sequences that still have the forward and reverse primers in the sequences.

-   The Raw Data is **demultiplexed**

-   A R1 and R2 fastq.gz file has been generated for each individual sample

-   All forward reads are binned into the R1 fastq.gz files

-   All reverse reads are binned into the R2 fastq.gz files

-   Other than demultiplexing; you can consider the Raw Data on BaseSpace as untouched (**The Forward and Reverse Primer Sequences have not been removed**)

Here I follow the QIIME2 Casava 1.8 paired-end demultiplexed fastq tutorial example on importing data, using the files provided to me by the Mr. DNA, Molecular Research.

## Data download

I got an email from Mr. DNA with a Dropbox link to the data files, where I downloaded two .zip folders; one had raw data files and the other had analysis pipeline files that Mr. DNA generated.

Here I am working with the raw data files located in `coral-pae-temp/analysis/microbiome/rawdata/demux`

In the `demux` folder is a `fastq.gz` file for The file name includes the sample identifier and should look like `4.Ea_S1_L001_R1_001.fastq.gz`. The underscore-separated fields in this file name are:

1.  the sample identifier,

2.  the barcode sequence or a barcode identifier,

3.  the lane number,

4.  the direction of the read (i.e. R1 or R2, because these are paired-end reads), and

5.  the set number.

```{r}
library(reticulate)
```

```{r, engine='python'}
conda activate qiime2-2023.5
```

## Get demultiplexed sequences

```{r, engine='bash'}

unzip 

```
