---
title: "Microbiome Sequence Analysis"
subtitle: "Setting Up The Workflow: Using QIIME 2 in an RStudio Server Environment"
author: "Sarah Tanja"
format: html
editor: visual
toc: true
toc-title: Contents <i class="bi bi-bookmark-heart"></i>
toc-depth: 5
toc-location: left
reference-location: margin
citation-location: margin
bibliography: ./references/qiime2.bib
---

## Overview

## Gitignore Large Files

In this project I'm dealing with large files that will end up in the `./rawdata` and `./output` directories. Any file larger than 100 MB cannot be pushed to GitHub (\*without paying for GitHub Large File Storage). To avoid that, I use two methods:

1.  Built-in hooks to automatically ignore files larger than 100 MB (no matter the directory or file name!)

-   Create a new text file in the `.git/hooks/` directory of your repository called `pre-commit`

-   *Select the `More` tab with the gear icon under the RStudio Files navigator bar and select 'show hidden files' to see the .git folder*.

-   Navigate to the `.git/hooks` folder

-   Rename `pre-commit.sample` file to `pre-commit`

-   Add the following text to the `.git/hooks/pre-commit` file:

```         
#!/bin/sh

# Maximum file size (in bytes)
max_file_size=104857600

# Find all files larger than max_file_size and add them to the .gitignore file
find . -type f -size +$max_file_size -exec echo "{}" >> .gitignore \;
```

-   Save the file changes

This code sets the max_file_size variable to 100 MB and then uses the find command to locate all files in the repository that are larger than the specified max_file_size. The exec option of the find command appends the name of each file that matches the criteria to the .gitignore file.

Save the pre-commit file and make it executable by running the following command in Terminal from your base git directory:

```{r, engine=bash}
cd ../../
chmod +x .git/hooks/pre-commit
```

With these changes, whenever you run a git commit command, Git will first execute the pre-commit hook, which will automatically add any files larger than 100 MB to the .gitignore file. This will prevent Git from tracking these files in the repository going forward.

2.  Manually execute code to find and add files greater than 100 MB to .gitignore before committing

Use the following code as a bash terminal command to find and add files greater than 100 MB from the repo to .gitignore

```{r, engine=bash}
cd ../../
find . -size +1G | sed 's|^./||g' | cat >> .gitignore
```

::: callout-tip
This handy code came from [this](https://github.com/sr320/course-fish546-2015/issues/43) Robert's Lab GitHub Issue
:::

## Install QIIME 2

Here I am installing QIIME 2 on the Roberts Lab `raven` server within a `conda` environment by following the steps outlined in the [QIIME 2 docs webpage](https://docs.qiime2.org/2023.5/install/).

> ### **Installing Miniconda**
>
> [Miniconda](https://docs.conda.io/en/latest/miniconda.html) provides the `conda` environment and package manager, and is the recommended way to install QIIME 2. Follow the [Miniconda instructions](https://conda.io/projects/conda/en/latest/user-guide/install/index.html) for downloading and installing Miniconda. You may choose either Miniconda2 or Miniconda3 (i.e. Miniconda Python 2 or 3). QIIME 2 will work with either version of Miniconda. It is important to follow all of the directions provided in the [Miniconda instructions](https://conda.io/projects/conda/en/latest/user-guide/install/index.html), particularly ensuring that you run `conda init` at the end of the installation process, to ensure that your Miniconda installation is fully installed and available for the following commands .

I choose to work with Miniconda3 and Python3 and am installing on a Linux OS (The Roberts Lab `raven` server).

In the following line of code I am using `curl -O` to download `miniconda3 Linux 64-bit` installer into an `applications` folder that exists outside of my repository and is the central place where I am downloading software.

#### Download the Miniconda installer for Linux

::: callout-tip
Instructions were followed from conda.io website \[here\](https://conda.io/projects/conda/en/latest/user-guide/install/linux.html)
:::

```{r, engine=bash, title='curl Linux Installer'}
cd ../../../../applications
curl -O https://repo.anaconda.com/miniconda/Miniconda3-py311_23.5.2-0-Linux-x86_64.sh
ls
```

#### Verify installer hashes

The SHA256 hash for Python 3.11 Miniconda3 Linux 64-bit is: `634d76df5e489c44ade4085552b97bebc786d49245ed1a830022b0b406de5817`

Download the installer file and before installing verify it as follows:

```{r, engine=bash, title='verify installer hash'}
cd ../../../../applications
sha256sum Miniconda3-py311_23.5.2-0-Linux-x86_64.sh
```

#### Run Miniconda

Copy the code below into the `Terminal` and press enter

```         
cd ../../../../applications
bash Miniconda3-py311_23.5.2-0-Linux-x86_64.sh
```

Follow all prompts in the Terminal.

#### Test installation

```{r, engine=bash}
conda list
```
